% Document class and font size
\documentclass[a4paper,11pt]{extarticle}

% Packages
\usepackage[utf8]{inputenc} % For input encoding
\usepackage{geometry} % For page margins
\geometry{a4paper, margin=0.75in} % Set paper size and margins
\usepackage{titlesec} % For section title formatting
\usepackage{enumitem} % For itemized list formatting
\usepackage{hyperref} % For hyperlinks

% Formatting
\setlist{noitemsep} % Removes item separation
\titleformat{\section}{\large\bfseries}{\thesection}{1em}{}[\titlerule] % Section title format
\titlespacing*{\section}{0pt}{\baselineskip}{\baselineskip} % Section title spacing

% Begin document
\begin{document}

% Disable page numbers
\pagestyle{empty}

% Header
\begin{center}
\textbf{\Large Zain Siddiqui}\\[2pt] % Name
St. Louis, MO| \href{mailto:example@example.com}{s.zain@wustl.edu} | 706-455-1501  % Contact info
\end{center}

\section{PERSONAL BACKGROUND}

Born December 16, 1995 in Cleveland, TN, USA. 

% Education Section
\section*{EDUCATION}
\noindent
\textbf{Washington University in St. Louis}, St. Louis, MO \hfill Enrolled: Fall 2021| Expected Graduation: Spring 2026\\ % University name and location
PhD in Mathematics\\
Thesis Advisor: Renato Feres

\noindent
\\
\textbf{Georgia Institute of Technology} \hfill Fall 2014-Spring 2019\\
BS in Mathematics
% Experience Section
\section*{RESEARCH INTERESTS}
\noindent
Some of my interests are dynamics, stochastic processes, and ergodic theory.\\

\noindent
In particular, I am interested in dynamical billiards. Specifically, I am focused on polygonal billiards and the topology of their unfoldings, and I wish to understand these unfoldings using both topological and stochastic methods.
% Projects Section
\section*{PUBLICATIONS}
\noindent
Don't rush me!

% Skills Section
\section*{TEACHING EXPERIENCE}

\textbf{Washington University in St. Louis}

\begin{itemize}
    \item Directed Reading Program, \emph{Geometry: Euclid and Beyond}
    \hfill Spring 2024
    \item Calculus I Teaching Assistant
    \hfill Spring 2024
    \item Calculus II Teaching Assistant \hfill Fall 2023
    \item Calculus II Teaching Assistant \hfill Spring 2023
\end{itemize}

\noindent
\textbf{Georgia Institute of Technology}

\begin{itemize}
    \item Calculus III Teaching Assistant \hfill Spring 2019
    \item Calculus III Teaching Assistant \hfill Fall 2018
    \item Calculus III Teaching Assistant \hfill Spring 2018
    \item Calculus III Teaching Assistant \hfill Fall 2017
    \item Calculus III Teaching Assistant \hfill Fall 2016
\end{itemize}

\section*{TALKS}
\begin{itemize}
    \item Szegő Seminar, Washington University in St. Louis, \emph{What an Incidence!: Hilbert's Axioms in Geometry} - March 2024
    \item Candidacy Talk, Washington University in St. Louis, \emph{Unfolding a Billiard Table} - March 2024
    
\end{itemize}


% End document
\end{document}
