\documentclass{article}

\usepackage{amsmath, amsthm, amssymb, amsfonts}
\usepackage{thmtools}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{geometry}
\usepackage{float}
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{framed}
\usepackage[dvipsnames]{xcolor}
\usepackage{tcolorbox}

\colorlet{LightGray}{White!90!Periwinkle}
\colorlet{LightOrange}{Orange!15}
\colorlet{LightGreen}{Green!15}

\newcommand{\HRule}[1]{\rule{\linewidth}{#1}}

\declaretheoremstyle[name=Theorem,]{thmsty}
\declaretheorem[style=thmsty,numberwithin=section]{theorem}
\tcolorboxenvironment{theorem}{colback=LightGray}

\declaretheoremstyle[name=Proposition,]{prosty}
\declaretheorem[style=prosty,numberlike=theorem]{proposition}
\tcolorboxenvironment{proposition}{colback=LightOrange}

\declaretheoremstyle[name=Principle,]{prcpsty}
\declaretheorem[style=prcpsty,numberlike=theorem]{principle}
\tcolorboxenvironment{principle}{colback=LightGreen}

\setstretch{1.2}
\geometry{
    textheight=9in,
    textwidth=5.5in,
    top=1in,
    headheight=12pt,
    headsep=25pt,
    footskip=30pt
}

% ------------------------------------------------------------------------------

\begin{document}

% ------------------------------------------------------------------------------
% Cover Page and ToC
% ------------------------------------------------------------------------------


% ------------------------------------------------------------------------------

\section*{Introduction}

First, I'd like to say that mathematics can be very difficult, and everyone 
has ``blind spots'' in their understanding. Don't put pressure on yourself
to know it all. Just do as much you reasonably can. 

Also, I want you to \textbf{ask questions}. This is an important part
of learning mathematics (and anything else, really). Asking a question 
is not some sign of weakness or a lack of intelligence. I want you to
ask a question even if it feels ``stupid''. Confusion is a normal state
of affairs in mathematics, and it can take a lot of time and effort to 
clear this confusion up. Quite often, you just need the right shift of perspective
to make sense of something, and different people have different ways
of understanding any one given thing. This is to say that an explanation that 
works very well for one person may be very unclear for another.




% ------------------------------------------------------------------------------

\end{document}